---
title: "Manuscript_Tables_LongCovid"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 6
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, cache = FALSE, cache.lazy = FALSE)
```

```{r include=FALSE}
# packages
library(dplyr)
library(ggplot2)
library(kableExtra)
library(stringr)
library(here)
library(cowplot)
library(lubridate)

Sys.setlocale("LC_ALL", "English")
options(knitr.kable.NA = " ")

# SIDIAP results
database_name <- "SIDIAP"
folder <- paste0("Results_", database_name)

files <-list.files(path= folder, pattern="*.csv") 
files <- str_remove(files, paste0(".csv"))

for (i in 1:length(files)){
  assign(
    files[i], 
    read.csv(paste0(folder, "/", files[i], ".csv"), header = TRUE, sep = ",")
  )
} 

col1 <- "#3A6D82"
col2 <- "#9F5339"

# CPRD results
database_name <- "CPRD"
folder <- paste0("Results_", database_name)

files <-list.files(path= folder, pattern="*.csv")
files <- str_remove(files, paste0(".csv"))


for (i in 1:length(files)){
  assign(
    files[i], 
    read.csv(paste0(folder, "/", files[i], ".csv"), header = TRUE, sep = ",")
  )
}

get_baseline_main_table <- function(sidiap_data, cprd_data) {
  Table <- sidiap_data %>%
    select(X, COVID.19.infection, Tested.negative..earliest) %>%
    full_join(cprd_data%>%
                select("X",
                       "COVID.infection.CPRD" = "COVID.19.infection"  ,
                       "Tested.negative.CPRD" = "Tested.negative..earliest"
                ) %>% 
          mutate(X = ifelse(X == "Trimester = jan-apr 2021 (%)",
                            "   Jan-apr 2021"  ,
                            ifelse(X == "Vaccination status = first dose vaccination (%)",                "   First dose vaccination", X)))) 
  Table1 <- kable(Table[c(1:34),],
                  col.names = c("", 
                                "COVID-19\ninfection", 
                                "First SARS-CoV-2 negative test ",
                                "COVID-19\ninfection", 
                                "First SARS-CoV-2 negative test "
                  )) %>%  
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  fixed_thead = T
    ) %>%
    add_header_above(c(" ", "SIDIAP" = 2, "CPRD" = 2)) %>%
    pack_rows("Comorbidities",26, 34) 
  Table1
}

# symptoms table
get_baseline_sup_table <- function(sidiap_data, cprd_data) {
  Table <- sidiap_data %>%
    select(-"COVID.19.infection",-"Tested.negative..earliest") %>%
    left_join(cprd_data%>%
                select("X",
                       "First.infection.CPRD" = "First.infection",
                       "Reinfections.CPRD" = "Reinfections",
                       "Tested.negative.all.CPRD" = "Tested.negative..all",
                ) %>% 
          mutate(X = ifelse(X == "Trimester = jan-apr 2021 (%)",
                            "   Jan-apr 2021"  ,
                     ifelse(X == "Vaccination status = first dose vaccination (%)",
                                   "   First dose vaccination", X)))) 
  Table1 <- kable(Table[c(1:34),],
                  col.names = c("",
                                "First infection",
                                "Re-infections",
                                "SARS-CoV-2 negative tests, all records",
                                "First infection",
                                "Re-infections",
                                "SARS-CoV-2 negative tests, all records"
                  ),
                  row.names = FALSE) %>%  
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  fixed_thead = T
    ) %>%
    add_header_above(c(" ", "SIDIAP" = 3, "CPRD" = 3)) %>%
    pack_rows("Comorbidities",26, 34) 
  Table1
}

get_matched_table_negative <- function(sidiap_data, cprd_data) {
  Table <- sidiap_data %>%
  left_join(cprd_data%>%
              select("X",
                     "COVID-19 infection, CPRD" = "New.infections"  ,
                     "Tested negative, CPRD" = "Tested.negative..earliest",
                     "SMD, CPRD" = "SMD"
              ) %>% 
          mutate(X = ifelse(X == "Trimester = jan-apr 2021 (%)",
                            "   Jan-apr 2021"  ,
                            ifelse(X == "Vaccination status = first dose vaccination (%)",
                                   "   First dose vaccination", X))))
Table1 <- kable(Table[c(1:34),],
                col.names = c("", "COVID-19\ninfection",
                              "First SARS-CoV-2 negative test ", "SMD", 
                              "COVID-19\ninfection",
                              "First SARS-CoV-2 negative test ", "SMD"
                )) %>%  
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                fixed_thead = T
  ) %>%
  add_header_above(c(" ", "SIDIAP" = 3, "CPRD" = 3)) %>%
  pack_rows("Comorbidities",26, 34) 
Table1
}

get_matched_table_first_reinfection <- function(sidiap_data, cprd_data) {
  Table <- sidiap_data %>%
    left_join(cprd_data%>%
                select("X",
                       "First infection, CPRD" = "First.infection"  ,
                       "Reinfections, CPRD" = "Reinfections",
                       "SMD, CPRD" = "SMD"
                )%>% 
          mutate(X = ifelse(X == "Trimester = jan-apr 2021 (%)",
                            "   Jan-apr 2021"  ,
                   ifelse(X == "Vaccination status = first dose vaccination (%)",
                                   "   First dose vaccination", X))))
  Table1 <- kable(Table[c(1:34),],
                  col.names = c("", "First\ninfection", 
                                "Re-infections", "SMD",
                                "First\ninfection",
                                "Re-infections", "SMD"
                  )) %>%  
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  fixed_thead = T
    ) %>%
    add_header_above(c(" ", "SIDIAP" = 3, "CPRD" = 3)) %>%
    pack_rows("Comorbidities",26, 34) 
  Table1
}

mytheme <-  theme_bw() +
  theme(
    #   text = element_text(family = "A"),
    panel.spacing = unit(1, "lines"),    
    axis.line.x.bottom = element_line(color = 'black'),
    axis.line.y.left   = element_line(color = 'black'),
    panel.border = element_blank(),
    panel.grid.major.y = element_blank() ,
    # explicitly set the horizontal lines (or they will disappear too)
    panel.grid.major.x = element_line( size=.1, color="#F5F4F6" ),
    legend.position = "top",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    axis.text.y = element_text(size = 14,),
    axis.text.x = element_text(size = 14,),
    axis.ticks.x =element_line(),
    axis.ticks.y =element_blank(),
    # strip.text.x.left = element_text(size = 14, face = "bold", angle = 90),
    strip.text.y = element_text(size = 14, face = "bold"),
    strip.background = element_blank(),
    #  panel.grid.major.y = element_blank() , # remove horizontal lines
    # top, rigth, bottom.left
    plot.margin = unit(c(0,0,1,0), "lines")
  )


# symptoms table
get_symptoms_main_table <- function(sidiap_data, cprd_data) {
  Table <- sidiap_data %>%
    left_join(cprd_data%>%
                select("X",
                       "COVID.infection.CPRD" = "COVID.19.infection"  ,
                       "First.infection.CPRD" = "First.infection",
                       "Reinfections.CPRD" = "Reinfections",
                       "Tested.negative.all.CPRD" = "Tested.negative..all",
                       "Tested.negative.CPRD" = "Tested.negative..earliest"
                ) %>% 
          mutate(X = ifelse(X == "Trimester = jan-apr 2021 (%)",
                            "   Jan-apr 2021"  ,
                     ifelse(X == "Vaccination status = first dose vaccination (%)",
                                   "   First dose vaccination", X)))) 
  Table1 <- kable(Table[c(1,35:74),],
                  col.names = c("",
                                "COVID-19\ninfection", 
                                "First infection",
                                "Re-infections",
                                "SARS-CoV-2 negative tests",
                                "First SARS-CoV-2 negative test ",
                                 "COVID-19\ninfection", 
                                "First infection",
                                "Re-infections",
                                "SARS-CoV-2 negative tests",
                                "First SARS-CoV-2 negative test "
                  ),
                  row.names = FALSE) %>%  
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  fixed_thead = T
    ) %>%
       pack_rows("Symptoms",3,27 ) %>%
    add_header_above(c("", "SIDIAP" = 5, "CPRD" = 5)) 
  Table1
}

get_rr_infection_negative_plot <- function(data){
plot <- data  %>%
    ggplot(aes(
      x = relative_risk,
      y = reorder(symptom, relative_risk),
      xmin = low_ci,
      xmax = up_ci,
      shape = database,
      color = database)) +
    geom_point(size = 2, position = position_dodge(width = 1)) +
    geom_errorbar(width = 0, size = 1, position = position_dodge(width = 1)) +
    scale_y_discrete(expand=c(0,0), position = "left") +
    scale_x_log10(breaks = c(0.5, 0.8, 1,2, 4, 6),
                  labels = c("0.5","0.8", "1","2", "4", "6"),
                  limits = c(0.5, 8)) +  
    scale_colour_manual(values = c(col1, col2)) +
    geom_vline(
      xintercept = 1, colour = "#000000",
      linetype = 1
    ) +
    mytheme +
    xlab("\nRR (with 95% CI)") +
    ylab("")  +
    guides(color = guide_legend(title = "Database"),
           shape = guide_legend(title = "Database")) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5, 9.5,10.5,11.5,
                     12.5,13.5,14.5,15.5,16.5,17.5,18.5,19.5,20.5,21.5,22.5,23.5),
                linetype = "dashed",
               color = "#D2D6DB"
    )
plot
}

get_rr_infection_negative_plot_facet <- function(data){
plot <- data  %>%
    ggplot(aes(
      x = relative_risk,
      y = reorder(symptom, relative_risk),
      xmin = low_ci,
      xmax = up_ci,
      shape = database,
      color = database)) +
    geom_point(size = 2, position = position_dodge(width = 1)) +
    geom_errorbar(width = 0, size = 1, position = position_dodge(width = 1)) +
    facet_grid(cols = vars(wave)) +
    scale_y_discrete(expand=c(0,0), position = "left") +
    scale_x_log10(breaks = c(0.5, 0.8, 1,2, 4, 6),
                  labels = c("0.5","0.8", "1","2", "4", "6"),
                  limits = c(0.5, 8)) +  
    scale_colour_manual(values = c(col1, col2)) +
    geom_vline(
      xintercept = 1, colour = "#000000",
      linetype = 1
    ) +
    mytheme +
    xlab("\nRR (with 95% CI)") +
    ylab("")  +
    guides(color = guide_legend(title = "Database"),
           shape = guide_legend(title = "Database")) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5, 9.5,10.5,11.5,
                     12.5,13.5,14.5,15.5,16.5,17.5,18.5,19.5,20.5,21.5,22.5,23.5),
                linetype = "dashed",
               color = "#D2D6DB"
    )
plot
}


get_rr_first_reinfection_plot <- function(data){
plot <- data  %>%
    ggplot(aes(
      x = relative_risk,
      y = reorder(symptom, -relative_risk),
      xmin = low_ci,
      xmax = up_ci,
      shape = database,
      color = database)) +
    geom_point(size = 2, position = position_dodge(width = 1)) +
    geom_errorbar(width = 0, size = 1, position = position_dodge(width = 1)) +
    scale_y_discrete(expand=c(0,0), position = "left") +
     scale_x_log10(breaks = c(0.6, 0.8, 1, 2),
                   labels = c("0.6", "0.8", "1", "2"),
                  limits = c(0.5, 2)
                  ) +  
    scale_colour_manual(values = c(col1, col2)) +
    geom_vline(
      xintercept = 1, colour = "#000000",
      linetype = 1
    ) +
    mytheme +
    xlab("\nRR (with 95% CI)") +
    ylab("")  +
    guides(color = guide_legend(title = "Database"),
           shape = guide_legend(title = "Database")) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5, 9.5,10.5,11.5,
                     12.5,13.5,14.5,15.5,16.5,17.5,18.5,19.5,20.5,21.5,22.5,23.5),
                linetype = "dashed",
               color = "#D2D6DB"
    )
plot
}

get_rr_first_reinfection_plot_facet <- function(data){
plot <- data  %>%
    ggplot(aes(
      x = relative_risk,
      y = reorder(symptom, -relative_risk),
      xmin = low_ci,
      xmax = up_ci,
      shape = database,
      color = database)) +
    geom_point(size = 2, position = position_dodge(width = 1)) +
    geom_errorbar(width = 0, size = 1, position = position_dodge(width = 1)) +
    facet_grid(cols = vars(wave)) +
    scale_y_discrete(expand=c(0,0), position = "left") +
     scale_x_log10(breaks = c(0.1, 0.2, 0.4, 0.6, 0.8, 1, 2),
                   labels = c("0.1", "0.2", "0.4","0.6", "0.8", "1", "2"),
                  limits = c(0.04, 2)
                  ) +  
    scale_colour_manual(values = c(col1, col2)) +
    geom_vline(
      xintercept = 1, colour = "#000000",
      linetype = 1
    ) +
    mytheme +
    xlab("\nRR (with 95% CI)") +
    ylab("")  +
    guides(color = guide_legend(title = "Database"),
           shape = guide_legend(title = "Database")) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5, 9.5,10.5,11.5,
                     12.5,13.5,14.5,15.5,16.5,17.5,18.5,19.5,20.5,21.5,22.5,23.5),
                linetype = "dashed",
               color = "#D2D6DB"
    )
plot
}

variables <- ls()
for (variable in variables) {
  x <- get(variable)
  if ("vaccine" %in% colnames(x)) {
    x <- x %>% 
      dplyr::filter(.data$vaccine == "overall") %>% 
      dplyr::select(-"vaccine")
    assign(variable, x)
  }
}

```

## Main results
### Table 1: Baseline characteristics of individuals with COVID-19 and tested negative (earliest record)
```{r echo=FALSE}
table1 <- get_baseline_main_table(sidiap_data = Table1_90days_SIDIAP,
                                  cprd_data = Table1_90days_CPRD)
table1
```

### Figure 1: Incidence rates of COVID-19 and long-COVID-19
```{r, fig.dim = c(8, 8), echo = FALSE}
IR_data <- rbind(Incidence_rates_Covid_LongCov_CPRD, Incidence_rates_Covid_LongCov_SIDIAP)
Proportion_data <- rbind(Proportion_LongCovid_90days_CPRD, Proportion_LongCovid_90days_SIDIAP)

data_IR_fig90 <- IR_data %>%
  filter(cohort != "Long COVID-19 (28 days)" ) %>%
  filter(!(is.na(n_events))) %>%
  mutate(sex= as.character(sex)) %>%
  mutate(incidence_start_date = as.Date(incidence_start_date)) %>%
  select(
    incidence_start_date, n_persons, n_events, person_years, 
    ir_100000_pys = incidence_100000_pys,  
    ir_100000_pys_95CI_lower = incidence_100000_pys_95CI_lower, 
    ir_100000_pys_95CI_upper = incidence_100000_pys_95CI_upper, 
    age_group, sex, cohort, database
  ) %>%
  rbind(Proportion_data %>%
          mutate(person_years = person_days/365) %>%
          select(-person_days) %>%
          filter(cohort=="COVID-19 infection") %>%
          mutate(cohort = "Long COVID-19 among those with COVID-19, 90 days")
  ) %>%
  mutate(age_group = ifelse(age_group == "18 to 150", "Overall", 
                            ifelse(age_group == "80 to 150", "\u226580", age_group))) %>%
  mutate(age_group = factor(age_group,
                            levels = c("Overall", "18 to 34", "35 to 49", "50 to 64", "65 to 79", "\u226580"))) %>%
      mutate(cohort = ifelse(cohort == "Long COVID-19 among those with COVID-19, 90 days", 
                           "Long COVID symptoms among those with COVID-19",
                           ifelse(cohort == "Long COVID-19 (90 days)", "Long COVID symptoms after COVID-19 among the general population", cohort))) %>%
  mutate(cohort = factor(cohort, levels = c(
    "COVID-19 infections",
     "Long COVID symptoms after COVID-19 among the general population",
     "Long COVID symptoms among those with COVID-19"
  )))


plot_IR_overall <- function(age_to_filter, data){ 
  plot <- data %>%
  filter(age_group %in% age_to_filter) %>%
  ggplot() +
#  geom_smooth(aes(x =incidence_start_date, y =  ir_100000_pys), method="auto", se=TRUE, fullrange=FALSE, level=0.95)+
geom_line(aes(x =incidence_start_date, y =  ir_100000_pys, color = database), size = 1.2)+
geom_ribbon(aes(x =incidence_start_date, y =  ir_100000_pys,
                ymin = ir_100000_pys_95CI_lower, ymax = ir_100000_pys_95CI_upper,
                group= database
                ), alpha = 0.3) +
scale_y_continuous(limits = c(0,NA)) +
scale_x_date(date_breaks = "2 month",
               date_labels = ("%b %y"),
               expand = c(0,0)) +
        scale_colour_manual(values = c(col1, col2)) +
  theme_bw()+
  theme(legend.position="top",
        plot.margin = unit(c(0.1, 0.5, 0.1, 0.1), "cm"),
        strip.text.y = element_text(angle = 0),
    strip.background =element_rect(fill="#F5F5F5")) +
  ylab("IR per 100,000 p-y") +
    facet_wrap(~cohort, scales = "free_y", ncol = 1) +
 # facet_grid(rows = vars(cohort),
 #            scales = "fixed") +
    xlab("Date of COVID-19 infection") +
        guides(color = guide_legend(title = "Database"), 
               fill = guide_legend(title = "Database"))
   plot         
  }


fig_IR_overall <- plot_IR_overall(age_to_filter = "Overall", data = data_IR_fig90) 
fig_IR_overall
```

### Figure 2A: RR of long COVID symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 90 days, overall
```{r, fig.dim = c(8, 8), echo = FALSE}
## Relative risks figure
 RR_data <- rbind(RR_new_covid_tested_negative_earliest_90days_SIDIAP,
                  RR_new_covid_tested_negative_earliest_90days_CPRD) %>%
  filter(wave == "overall") %>%
  select(-"wave") %>%
  mutate(symptom = ifelse(symptom == "Altered smell taste", "Altered smell or taste",
                          ifelse(symptom =="Cognitive dysfunction brain fog", "Cognitive dysfunction",
                                 ifelse(symptom == "Fatigue malaise", "Fatigue or malaise",
                                        ifelse(symptom == "Muscle spams pain", "Muscle spasms or pain", symptom
                                        )))))

gg.rr.covid_negative_both_90 <- get_rr_infection_negative_plot(data = RR_data)
gg.rr.covid_negative_both_90 
 
```

```{r, fig.dim = c(8, 8), echo = FALSE}
gg.rr.covid_negative_both <-  gg.rr.covid_negative_both_90 + 
   theme(axis.text.x = element_text(size = 14, angle = 90,hjust=0.95,vjust=0.2)) +
  coord_flip()
gg.rr.covid_negative_both
```

### Figure 2B: RR of long COVID symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 90 days, by waves

```{r, fig.dim = c(24, 8), echo = FALSE}
## Relative risks figure
 RR_data <- rbind(RR_new_covid_tested_negative_earliest_90days_SIDIAP,
                  RR_new_covid_tested_negative_earliest_90days_CPRD) %>%
  filter(wave != "overall") %>%
  mutate(symptom = ifelse(symptom == "Altered smell taste", "Altered smell or taste",
                          ifelse(symptom =="Cognitive dysfunction brain fog", "Cognitive dysfunction",
                                 ifelse(symptom == "Fatigue malaise", "Fatigue or malaise",
                                        ifelse(symptom == "Muscle spams pain", "Muscle spasms or pain", symptom
                                        ))))) %>%
  mutate(wave = factor(wave, levels = c("wild", "alpha", "delta", "omicron")))

gg.rr.covid_negative_both_90 <- get_rr_infection_negative_plot_facet(data = RR_data)
gg.rr.covid_negative_both_90 
 
```

```{r, fig.dim = c(8, 8), echo = FALSE}
gg.rr.covid_negative_both <-  gg.rr.covid_negative_both_90 + 
   theme(axis.text.x = element_text(size = 14, angle = 90,hjust=0.95,vjust=0.2)) +
  coord_flip()
gg.rr.covid_negative_both
```


### Figure 3: RR of long-COVID symptoms: first infection vs re-infection - 90 days
```{r, fig.dim = c(8, 8), echo = FALSE}
RR_data <- rbind(RR_first_infection_reinfections_90days_SIDIAP,
                  RR_first_infection_reinfections_90days_CPRD) %>%
  filter(wave == "overall") %>%
  select(-"wave") %>%
  mutate(symptom = ifelse(symptom == "Altered smell taste",
                          "Altered smell or taste",
                    ifelse(symptom =="Cognitive dysfunction brain fog",
                           "Cognitive dysfunction",
                    ifelse(symptom == "Fatigue malaise",
                           "Fatigue or malaise",
                    ifelse(symptom == "Muscle spams pain",
                           "Muscle spasms or pain", symptom
                                        )))))


 gg.rr.first_reinfection_90 <-  get_rr_first_reinfection_plot(data = RR_data)
gg.rr.first_reinfection_90
```

## Supplementary
### Baseline characteristics and symptoms distribution
#### 1. Table: Baseline characteristics of first infections, re-infections and all negative SARS-CoV-2 tests records
```{r}
table1_sup <- get_baseline_sup_table(sidiap_data = Table1_90days_SIDIAP,
                                  cprd_data = Table1_90days_CPRD)
table1_sup
```

#### 2. Table: Symptoms distribution across cohorts (unmatched) - 90 days
```{r}
get_symptoms_main_table(sidiap_data = Table1_90days_SIDIAP, 
                        cprd_data = Table1_90days_CPRD)
```

#### 3. Table: Symptoms distribution across cohorts (unmatched) - 28 days
```{r}
get_symptoms_main_table(sidiap_data = Table1_28days_SIDIAP, 
                        cprd_data = Table1_28days_CPRD)
```

#### 4. Table: Baseline characteristics - COVID-19 infection vs first negative test matched 3:1
```{r echo=FALSE}
Sup_table_matched_negative_90 <- get_matched_table_negative(
  sidiap_data = Table_Matched_Infection_Tested_Negative_earliest_90days_SIDIAP,
  cprd_data = Table_Matched_Infection_Tested_Negative_earliest_90days_CPRD)
Sup_table_matched_negative_90 
```

#### 5. Table: Baseline characteristics: first COVID-19 infection vs re-infections matched 1:3
```{r echo=FALSE}
Sup_table_matched_first_reinfection_90 <- get_matched_table_first_reinfection(
  sidiap_data = Table_Matched_First_infection_Reinfections_90days_SIDIAP,
  cprd_data = Table_Matched_First_infection_Reinfections_90days_CPRD)
Sup_table_matched_first_reinfection_90 
```

### Incidence Rates
#### 1. Table: Incidence Rates of COVID-19 & long COVID symptoms
```{r}

data_table_IR <- data_IR_fig90 %>% 
  filter(age_group == "Overall" & sex == "Both") %>%
  relocate(database, cohort, incidence_start_date, n_events,
           person_years) %>%
  select(-n_persons, -age_group, -sex) %>%
  mutate(person_years = sprintf(person_years, fmt = '%.0f'),
    ir_100000_pys = sprintf(ir_100000_pys, fmt = '%.1f'),
         ir_100000_pys_95CI_lower = sprintf(ir_100000_pys_95CI_lower, fmt = '%.1f'),
         ir_100000_pys_95CI_upper = sprintf(ir_100000_pys_95CI_upper, fmt = '%.1f'),
         ci = paste0(ir_100000_pys, " [",
                     ir_100000_pys_95CI_lower, "-", ir_100000_pys_95CI_upper, "]")
         ) %>%
  select(-ir_100000_pys, -ir_100000_pys_95CI_lower, -ir_100000_pys_95CI_upper)

Table_IR <-  kable(data_table_IR,
                  col.names = c(
                                "Database",
                                "Cohort",
                                "COVID-19 test date",
                                "Number of events",
                                "Person-years",
                                "Incidence Rate"
                  ),
                  row.names = FALSE) %>%  
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  fixed_thead = T
    ) %>%
    collapse_rows(columns = c(1,2))

Table_IR
```

#### 2. Figure: Incidence Rates by Sex and Age group
##### A. SIDIAP
```{r echo=FALSE}

brk_vec <- seq.Date(as.Date("2020-10-01"),
                    as.Date("2021-10-01"),by="2 month")

plot_IR <- function(age_to_filter, data, brk_vec){ 
  plot <- data %>%
  filter(age_group %in% age_to_filter) %>%
    mutate(cohort = ifelse(cohort == "COVID-19 infections", "COVID-19 infections",
                           ifelse(cohort == "Long COVID symptoms after COVID-19 among the general population", "Long COVID among\nthe general population", "Long COVID among\nthose infected"))) %>%
  ggplot() +
#  geom_smooth(aes(x =incidence_start_date, y =  ir_100000_pys), method="auto", se=TRUE, fullrange=FALSE, level=0.95)+
geom_line(aes(x =incidence_start_date,
              y =  ir_100000_pys,
              color = age_group), size = 1.2)+
#    geom_ribbon(aes(x =incidence_start_date, 
    # y =  ir_100000_pys, color = age_group,
    # ymin = ir_100000_pys_95CI_lower, ymax = ir_100000_pys_95CI_upper
    # ), alpha = 0.1) +
#scale_y_continuous(limits = c(0,NA)) +
scale_x_date(breaks = brk_vec,
 #date_breaks = "2 month",
               date_labels = ("%b %y"),
               expand = c(0,0)) +
scale_colour_manual(values = c("#90e0ef","#00b4d8","#0077b6","#023E8A", "#01022F"
  #"#000000","#0075c4","#5b8c5a","#d8973c","#dc392e","#d8973c","#dc392e"
  # "#bee9e8","#62b6cb",","#cae9ff","#5fa8d3"
                                     )) +
    theme_bw()+
    theme(
      #panel.spacing.x = unit(4, "mm"),
      strip.text.y.left =  element_text(angle = 0),
      legend.position="top",
      plot.margin = unit(c(0, 0, 0, 0), "cm"))+
  ylab("IR per 100,000 p-y") +
 facet_grid(cohort ~sex  , 
            scales = "free") +    
    xlab("Date of COVID-19 infection") +
        guides(color = guide_legend(title = "Age group"))
   plot         
  }


fig_IR_age_sex_SIDIAP  <- plot_IR(data = data_IR_fig90 %>% filter(database == "SIDIAP"),
                                     age_to_filter = c("18 to 34",
                                             "35 to 49",
                                             "50 to 64",
                                             "65 to 79",
                                             "\u226580"),
                                  brk_vec = brk_vec)
fig_IR_age_sex_SIDIAP 
```

```{r include=FALSE}

ggsave(paste0(here(),"/Results/Figures/IR_Sex_Age_SIDIAP.tiff"),
fig_IR_age_sex_SIDIAP ,
       dpi = 300,
       width = 9, height = 6)
```

##### B. CPRD
```{r echo=FALSE}

brk_vec <- seq.Date(as.Date("2020-10-01"),
                    as.Date("2020-12-01"),by="1 month")


fig_IR_age_sex_CPRD <-  plot_IR(data = data_IR_fig90 %>% filter(database == "CPRD AURUM"),
                                     age_to_filter = c(
                                       "18 to 34",
                                             "35 to 49",
                                             "50 to 64",
                                             "65 to 79",
                                             "\u226580"),
                                brk_vec = brk_vec)
fig_IR_age_sex_CPRD

ggsave(paste0(here(),"/Results/Figures/IR_Sex_Age_CPRD.tiff"),
fig_IR_age_sex_CPRD,
       dpi = 300,
       width = 9, height = 6)
```

### Relative Risks - infections vs first negative test 
#### 1. Table: RR of long COVID symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 90 days
```{r}
get_RR <- function(sidiap_data, cprd_data) {
  RR_infections <- rbind(sidiap_data, cprd_data
                      ) %>%
    filter(wave == "overall") %>%
    select(-"wave") %>%
  mutate(name_1 = "COVID-19 infection") %>%
  mutate(relative_risk = paste0(sprintf(relative_risk, fmt = '%.2f'),
                                " [",
                                sprintf(low_ci, fmt = '%.2f'),
                                ";",
                                sprintf(up_ci, fmt = '%.2f'),
                                "]"
         )) %>%
    select (-prop_1, -prop_2, -low_ci, -up_ci, 
                        -name_1, -name_2, -total_pop_1, -total_pop_2,) %>%
    relocate(database, .before = symptom)  %>%
    arrange(desc(database), desc(relative_risk))


table <- kable(RR_infections,  col.names = c(
     "Database",
                               "Symptom",
                              "Events COVID-19 infection",
                              "Events first negative test",
                             "Relative risk [with 95%CI]"
                             
                  ),
                  row.names = FALSE) %>%  
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  fixed_thead = T
    ) %>%
    collapse_rows(columns = 1)
table
}

table_RR_90 <- get_RR(sidiap_data = RR_new_covid_tested_negative_earliest_90days_SIDIAP,
                  cprd_data = RR_new_covid_tested_negative_earliest_90days_CPRD)
table_RR_90
```


#### 2. Table: RR of long COVID symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 28 days
```{r}
table_RR_28 <- get_RR(sidiap_data = RR_new_covid_tested_negative_earliest_28days_SIDIAP,
                  cprd_data = RR_new_covid_tested_negative_earliest_28days_CPRD)
table_RR_28
```

#### 3. Figure: RR of long COVID symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 28 days
```{r, fig.dim = c(8, 8), echo = FALSE}
## Relative risks figure
 RR_data <- rbind(RR_new_covid_tested_negative_earliest_28days_SIDIAP,
                  RR_new_covid_tested_negative_earliest_28days_CPRD) %>%
  filter(wave == "overall") %>%
  select(-"wave") %>%
  mutate(symptom = ifelse(symptom == "Altered smell taste", "Altered smell or taste",
                          ifelse(symptom =="Cognitive dysfunction brain fog", "Cognitive dysfunction",
                                 ifelse(symptom == "Fatigue malaise", "Fatigue or malaise",
                                        ifelse(symptom == "Muscle spams pain", "Muscle spasms or pain", symptom
                                        )))))

gg.rr.covid_negative_both_28 <- get_rr_infection_negative_plot(data = RR_data)
gg.rr.covid_negative_both_28 
```

### Relative Risks - re-infections vs first infection
#### 1. Table: RR of long-COVID symptoms, first infection vs re-infection - 90 days
```{r}
get_RR_numbers_re_infec <- function(sidiap_data, cprd_data) {
  RR_infections <- rbind(sidiap_data, cprd_data
                      ) %>%
    filter(wave == "overall") %>%
    select(-"wave") %>%
  mutate(relative_risk = paste0(sprintf(relative_risk, fmt = '%.2f'),
                                " [",
                                sprintf(low_ci, fmt = '%.2f'),
                                ";",
                                sprintf(up_ci, fmt = '%.2f'),
                                "]"
         )) %>%
    select (-prop_1, -prop_2, -low_ci, -up_ci, 
                        -name_1, -name_2, -total_pop_1, -total_pop_2,) %>%
    relocate(database,.before = symptom)  %>%
    arrange(desc(database),relative_risk)

table_RR_re <- kable(RR_infections,  col.names = c(
     "Database",
                               "Symptom",
                              "Events re-infection",
                              "Events first infection",
                             "Relative risk [with 95%CI]"
                             
                  ),
                  row.names = FALSE) %>%  
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  fixed_thead = T
    ) %>%
    collapse_rows(columns = 1)
table_RR_re
  
  
}

table_RR_re_90 <- get_RR_numbers_re_infec(sidiap_data = RR_first_infection_reinfections_90days_SIDIAP,
                                       cprd_data = RR_first_infection_reinfections_90days_CPRD)
 table_RR_re_90 
```

#### 2. Table: RR of long-COVID symptoms, first infection vs re-infection - 28 days
```{r}
table_RR_re_28 <- get_RR_numbers_re_infec(sidiap_data = RR_first_infection_reinfections_28days_SIDIAP,
                                       cprd_data = RR_first_infection_reinfections_28days_CPRD)
 table_RR_re_28
```

#### 3. Figure: RR of long-COVID symptoms, first infection vs re-infection - 28 days
```{r, fig.dim = c(8, 8), echo = FALSE}

RR_data <- rbind(RR_first_infection_reinfections_28days_SIDIAP,   RR_first_infection_reinfections_28days_CPRD) %>%
  filter(wave == "overall") %>% select(-"wave") %>%
  mutate(symptom = ifelse(symptom == "Altered smell taste",
                          "Altered smell or taste",
                    ifelse(symptom =="Cognitive dysfunction brain fog",
                           "Cognitive dysfunction",
                    ifelse(symptom == "Fatigue malaise",
                           "Fatigue or malaise",
                    ifelse(symptom == "Muscle spams pain",
                           "Muscle spasms or pain", symptom
                                        ))))) 


gg.rr.first_reinfection_28 <-  get_rr_first_reinfection_plot(data = RR_data) +  
  scale_x_log10(breaks = c(0.1, 0.2, 0.4, 0.6, 0.8, 1,2, 3),
                   labels = c("0.1", "0.2", "0.4","0.6", "", "1", "2", "3"),
                  limits = c(0.04,3))
                   
gg.rr.first_reinfection_28

```

### Relative Risks - infections vs all negative tests
#### 1. Table: RR of long COVID symptoms, COVID-19 infection vs all SARS-CoV-2 negative tests - 90 days
```{r}
table_RR_90 <- get_RR(sidiap_data = RR_new_covid_tested_negative_earliest_90days_SIDIAP,
                  cprd_data = RR_new_covid_tested_negative_earliest_90days_CPRD)
table_RR_90
```

#### 2. Figure: RR of long COVID symptoms, COVID-19 infection vs SARS-CoV-2 negative tests (all records) - 90 days
```{r, fig.dim = c(8, 8), echo = FALSE}

RR_data <- rbind(RR_new_covid_tested_negative_all_90days_SIDIAP,
                 RR_new_covid_tested_negative_all_90days_CPRD) %>%
  filter(wave == "overall") %>%
  select(-"wave") %>%
  mutate(symptom = ifelse(symptom == "Altered smell taste",
                          "Altered smell or taste",
                    ifelse(symptom =="Cognitive dysfunction brain fog",
                           "Cognitive dysfunction",
                    ifelse(symptom == "Fatigue malaise",
                           "Fatigue or malaise",
                    ifelse(symptom == "Muscle spams pain",
                           "Muscle spasms or pain", symptom
                                        )))))


gg.rr.covid_negative_all_90 <- get_rr_infection_negative_plot(data = RR_data)
gg.rr.covid_negative_all_90
```

#### 3. Table: RR of long COVID symptoms, COVID-19 infection vs all SARS-CoV-2 negative tests - 28 days
```{r}
table_RR_28 <- get_RR(sidiap_data = RR_new_covid_tested_negative_earliest_28days_SIDIAP,
                  cprd_data = RR_new_covid_tested_negative_earliest_28days_CPRD)
table_RR_28
```

#### 4. Figure: RR of long COVID symptoms, COVID-19 infection vs SARS-CoV-2 negative tests (all records) - 28 days
```{r, fig.dim = c(8, 8), echo = FALSE}

RR_data <- rbind(RR_new_covid_tested_negative_all_28days_SIDIAP,
                 RR_new_covid_tested_negative_all_28days_CPRD) %>%
  filter(wave == "overall") %>%
  select(-"wave") %>%
  mutate(symptom = ifelse(symptom == "Altered smell taste",
                          "Altered smell or taste",
                    ifelse(symptom =="Cognitive dysfunction brain fog",
                           "Cognitive dysfunction",
                    ifelse(symptom == "Fatigue malaise",
                           "Fatigue or malaise",
                    ifelse(symptom == "Muscle spams pain",
                           "Muscle spasms or pain", symptom
                                        )))))


gg.rr.covid_negative_all_28 <- get_rr_infection_negative_plot(data = RR_data)
gg.rr.covid_negative_all_28
```
